StreamReduce: Processing Data Streams\\and Avoiding Work

### 2012-05-11 paper/paper.tex
Run through spellchecker

With terabytes of data, however, it is scarcely necessary to examine
every entry to uncover trends, especially when one only cares about the most prominent
trends or results.
How does one decide which data to process then? Ultimately, this is problem dependent;
however, selctively deciding which data to
log is problematic because it can systemically mask some of the trends one would
try to uncover through analysis. In this paper, I introduce StreamReduce, a distributed system
for processing data streams. With StreamReduce I explore:
\begin{itemize}
  \item
    Randomly discarding data while trying to maintain a target accuracy level.
  \item
    Processing data using a variety of computation routines, some of which are less
    accurate than others.
\end{itemize}
Combining both of these techniques allows the cluster to save time by executing faster,
save resources by avoiding the need to process every datum, and minimize the impact of faults
in the execution of a job by treating failed tasks as discarded tasks.

@@

The goal of StreamReduce is to provide a MapReduce-style framework for processing streams. Such a
framework should include the ability to kill or discard tasks, be capable of
providing workers with serveral different computation routines to choose among, and pass data
between nodes using non-blocking routines.

@@

The Hadoop Online Prototype~\cite{Condie:EECS-2009-136} adds streaming to the MapReduce framework
by removing the materialize phase between mappers and reducers and between phases of a
MapReduce pipeline. Results are streamed from mappers to reducers as they become available.
At any given point, each reducer maintains an in-progress snapshot of the MapReduce job's
progress up to that point. This removes the synchronization barriers inherient in MapReduce
and allows incremental progress to be observed.

@@

The \emph{Data Stream Model}~\cite{Babcock:2002:MID:543613.543615} is a refinement of the Chronicle
Data Model. It has the following properties: stream items may arrive out of order,
streams are potentially unbounded in size, and elements in the stream are discared after they are
processed. This is the data model that StreamReduce uses.

@@

The most important piece Dynamic Knobs offers to the DSMS space is dynamic
\emph{peak load provisioning}. Their system can adjust downward the amount of resources,
machines, etc. required for normal load and quickly adjust these levels back up to
handle load spikes. Combined with Aurora's load shedding, peak load provisioning can
be used to create a DSMS that uses less resources in both the normal and peak load
states.

Accuracy-aware transformations~\cite{Zhu:2012:RAP:2103656.2103710} are a class of transformations
that alter a computation according to a set of probabilstic weights and an accuracy
target such that the computation performs more efficiently. StreamReduce uses the subclass
of transformations called \emph{substitution transformations} in worker nodes to adjust
the runtime characteristics of per-datum computation.

@@

\section{Design}
\label{sec:design}
StreamReduce is a distributed system for processing streams that can choose to kill tasks or
limit the amount of work it performs per task. Its design is modular and there are five
different types of nodes in a StreamReduce cluster:
\begin{itemize}
  \item
    \textbf{Spout}: A node external to the cluster that is a source of streaming data.
  \item
    \textbf{Fetcher}: Communicates with spouts to provide data for workers.
  \item
    \textbf{Worker}: Processes data returned by fetchers. They are the source of parallel computation
    in the cluster.
  \item
    \textbf{Collector}: Aggregates results from all workers and makes the result accessible via an API.
  \item
    \textbf{Master}: Manages the cluster by bringing nodes online to complete jobs. It is reponsible
    for all internode communication and storing to-be-processed data in several queues.
\end{itemize}
Figure~\ref{fig:systemdiagram} shows a system diagram of StreamReduce.

@@

Because one of StreamReduce's goals is to process (potentially infinite) streams, workers do not
save their progress to disk and are forgetful; once the master queries a worker for its
progress and the worker successfully returns it, the worker clears all of its state and
continues processing. As a consequence, workers are easy to add or remove from the worker
pool because they don't need to be initialized with long-term job state. Additionally, as
long as the master polls workers frequently enough, only a small
fraction of the total work is lost upon worker failure.
The master queries workers every 100ms by default.
\subsection{Collector}
Collector nodes aggregate the results of all workers. Each collector acts as a reducer would
in the MapReduce framework if MapReduce used a single reducer for all of the mappers.
Collectors differ from MapReduce reducers in that they receive their data piecewise, as it
is produced from workers, as opposed to after all mappers have completed. This limits the
types of stateless aggregation collectors can do to online operations. (Of course, a collector
can choose to retain copies of individual worker results and re-reduce them each time more
results are receieved, but this is not in the spirit of the framework.)

@@

StreamReduce sends the source of the Jobfile to every node participating in the job and uses
\texttt{eval} to
obtain a local instance of the Jobfile class. An example jobfile can be found at~\cite{github}.

The use of \texttt{eval} to instantiate Jobfile objects and the implementation language's (ruby)
ability to monkeypatch classes means that jobs can be modified on the fly just by sending
an updated Jobfile to job participants. For example, a job can be made lower fidelity in
the face of an impending time constraint, be made higher fidelity if more workers are added to
the cluster, or gracefully kill all fetchers which allows the job to end after all of the
various job queues are empty.
\subsection{Fault Tolerance}
StreamReduce can sustain the failure of any node other than the master. Because the master is responsible
for shuffling data between fetchers, workers, and collectors, a job cannot make progress without
the master.

@@

\section{Methodology}
\label{sec:methodology}
The cluster I used to test StreamReduce is a 5-node network consisting of~1 spout,~1 fetcher,~4 workers,~1
collector, and~1 master. Roles were partitioned among nodes as shown in
Figure~\ref{fig:clusterDiagram}. Worker nodes were XVM~\cite{sipb:xvm} VMs provisioned with~128MB RAM each.
The master, spout, fetcher, and collector ran on a mid-2010 Macbook Air laptop with~4GB of RAM. I configured
the cluster to kill tasks on the master node, before they were shipped to workers.

@@

The results of the word count job were then used in a script to compute the top~100 words.
The cluster and word count job could be used in a two-stage top-k job. This script filters
out SEO stopwords~\cite{stopwords} to achieve more interesting results.

@@

\begin{figure}
\includegraphics[width=\linewidth]{long-tail-ranks.png}
\caption{The frequencies of words in the Wikipedia Featured Articles corpus have a long-tailed distribution. This graph has had stopwords~\cite{stopwords} filtered out of it.}
\label{fig:wordDist}
\end{figure}

@@

\subsection{Speed vs. Accuracy Tradeoffs}
\begin{center}
\begin{table*}
  \begin{tabularx}{\linewidth}{|lr|lr|lr|}
    \hline
    \multicolumn{2}{|X|}{Work level = 1.0 and task kill rate = 0.0} &
    \multicolumn{2}{|X|}{Work level = 0.8 and task kill rate = 0.4} &
    \multicolumn{2}{|X|}{Work level = 0.4 and task kill rate = 0.8} \\ \hline
    33225 & title & 21184 & title & 5982 & title \\ \hline
    29356 & web & 18398 & web & 4689 & web \\ \hline
    23095 & book & 13591 & journal & 4225 & journal \\ \hline
    21821 & journal & 13429 & book & 3629 & year \\ \hline
    20124 & year & 12770 & publisher & 3617 & publisher \\ \hline
    19786 & publisher & 12550 & url & 3444 & url \\ \hline
    19478 & url & 12220 & year & 3410 & book \\ \hline
    17360 & news & 11511 & news & 3273 & news \\ \hline
    16685 & accessdate & 11082 & united & 2968 & accessdate \\ \hline
    16401 & united & 10691 & accessdate & 2582 & world \\ \hline
  \end{tabularx}
  \caption{Example top~10 results and their frequencies from several different job configurations. The top~10 do not change much one changes the paramters to make the job less accurate.}
  \label{table:top10}
\end{table*}
\end{center}

>>>>

With terabytes of data, however, it is scarcely necessary to examine
every entry to uncover trends, especially when one only cares about the most prominent
trends or results.
How does one decide which data to process then? Ultimately, this is problem dependent;
however, selectively deciding which data to
log is problematic because it can systemically mask some of the trends one would
try to uncover through analysis. In this paper, I introduce StreamReduce, a distributed
system
for processing data streams. With StreamReduce I explore:
\begin{itemize}
  \item
    Randomly discarding data while trying to maintain a target accuracy level.
  \item
    Processing data using a variety of computation routines, some of which are less
    accurate than others.
\end{itemize}
Combining both of these techniques allows the cluster to save time by executing faster,
save resources by avoiding the need to process every datum, and minimize the impact of
faults
in the execution of a job by treating failed tasks as discarded tasks.

@@

The goal of StreamReduce is to provide a MapReduce-style framework for processing
streams. Such a
framework should include the ability to kill or discard tasks, be capable of
providing workers with several different computation routines to choose among, and pass
data
between nodes using non-blocking routines.

@@

The Hadoop Online Prototype~\cite{Condie:EECS-2009-136} adds streaming to the
MapReduce framework
by removing the materialize phase between mappers and reducers and between phases of a
MapReduce pipeline. Results are streamed from mappers to reducers as they become
available.
At any given point, each reducer maintains an in-progress snapshot of the MapReduce job's
progress up to that point. This removes the synchronization barriers inherent in MapReduce
and allows incremental progress to be observed.

@@

The \emph{Data Stream Model}~\cite{Babcock:2002:MID:543613.543615} is a refinement of
the Chronicle
Data Model. It has the following properties: stream items may arrive out of order,
streams are potentially unbounded in size, and elements in the stream are discarded after
they are
processed. This is the data model that StreamReduce uses.

@@

The most important piece Dynamic Knobs offers to the DSMS space is dynamic
\emph{peak load provisioning}. Their system can adjust downward the amount of resources,
machines, etc\. required for normal load and quickly adjust these levels back up to
handle load spikes. Combined with Aurora's load shedding, peak load provisioning can
be used to create a DSMS that uses less resources in both the normal and peak load
states.

Accuracy-aware transformations~\cite{Zhu:2012:RAP:2103656.2103710} are a class of
transformations
that alter a computation according to a set of probabilistic weights and an accuracy
target such that the computation performs more efficiently. StreamReduce uses the subclass
of transformations called \emph{substitution transformations} in worker nodes to adjust
the runtime characteristics of per-datum computation.

@@

\section{Design}
\label{sec:design}
StreamReduce is a distributed system for processing streams that can choose to kill tasks or
limit the amount of work it performs per task. Its design is modular and there are five
different types of nodes in a StreamReduce cluster:
\begin{itemize}
  \item
    \textbf{Spout}: A node external to the cluster that is a source of streaming data.
  \item
    \textbf{Fetcher}: Communicates with spouts to provide data for workers.
  \item
    \textbf{Worker}: Processes data returned by fetchers. They are the source of parallel
computation
    in the cluster.
  \item
    \textbf{Collector}: Aggregates results from all workers and makes the result accessible via
an API.
  \item
    \textbf{Master}: Manages the cluster by bringing nodes online to complete jobs. It is
responsible
    for all internode communication and storing to-be-processed data in several queues.
\end{itemize}
Figure~\ref{fig:systemdiagram} shows a system diagram of StreamReduce.

@@

Because one of StreamReduce's goals is to process (potentially infinite) streams, workers do
not
save their progress to disk and are forgetful; once the master queries a worker for its
progress and the worker successfully returns it, the worker clears all of its state and
continues processing. As a consequence, workers are easy to add or remove from the worker
pool because they don't need to be initialized with long-term job state. Additionally, as
long as the master polls workers frequently enough, only a small
fraction of the total work is lost upon worker failure.
The master queries workers every 100ms by default.
\subsection{Collector}
Collector nodes aggregate the results of all workers. Each collector acts as a reducer would
in the MapReduce framework if MapReduce used a single reducer for all of the mappers.
Collectors differ from MapReduce reducers in that they receive their data piecewise, as it
is produced from workers, as opposed to after all mappers have completed. This limits the
types of stateless aggregation collectors can do to online operations. (Of course, a collector
can choose to retain copies of individual worker results and re-reduce them each time more
results are received, but this is not in the spirit of the framework.)

@@

StreamReduce sends the source of the Jobfile to every node participating in the job and uses
\texttt{eval} to
obtain a local instance of the Jobfile class. An example Jobfile can be found at~\cite{github}.

The use of \texttt{eval} to instantiate Jobfile objects and the implementation language's
(ruby)
ability to monkey-patch classes means that jobs can be modified on the fly just by sending
an updated Jobfile to job participants. For example, a job can be made lower fidelity in
the face of an impending time constraint, be made higher fidelity if more workers are added
to
the cluster, or gracefully kill all fetchers which allows the job to end after all of the
various job queues are empty.
\subsection{Fault Tolerance}
StreamReduce can sustain the failure of any node other than the master. Because the master
is responsible
for shuffling data between fetchers, workers, and collectors, a job cannot make progress
without
the master.

@@

\section{Methodology}
\label{sec:methodology}
The cluster I used to test StreamReduce is a 5-node network consisting of~1 spout,~1
fetcher,~4 workers,~1
collector, and~1 master. Roles were partitioned among nodes as shown in
Figure~\ref{fig:clusterDiagram}. Worker nodes were XVM~\cite{sipb:xvm} VMs provisioned
with~128MB RAM each.
The master, spout, fetcher, and collector ran on a mid-2010 Macbook Air laptop with~4GB of
RAM\@. I configured
the cluster to kill tasks on the master node, before they were shipped to workers.

@@

The results of the word count job were then used in a script to compute the top~100 words.
The cluster and word count job could be used in a two-stage top-k job. This script filters
out SEO stop words~\cite{stopwords} to achieve more interesting results.

@@

\begin{figure}
\includegraphics[width=\linewidth]{long-tail-ranks.png}
\caption{The frequencies of words in the Wikipedia Featured Articles corpus have a long-
tailed distribution. This graph has had stop words~\cite{stopwords} filtered out of it.}
\label{fig:wordDist}
\end{figure}

@@

\subsection{Speed vs. Accuracy Tradeoffs}
\begin{center}
\begin{table*}
  \begin{tabularx}{\linewidth}{|lr|lr|lr|}
    \hline
    \multicolumn{2}{|X|}{Work level = 1.0 and task kill rate = 0.0} &
    \multicolumn{2}{|X|}{Work level = 0.8 and task kill rate = 0.4} &
    \multicolumn{2}{|X|}{Work level = 0.4 and task kill rate = 0.8} \\ \hline
    33225 & title & 21184 & title & 5982 & title \\ \hline
    29356 & web & 18398 & web & 4689 & web \\ \hline
    23095 & book & 13591 & journal & 4225 & journal \\ \hline
    21821 & journal & 13429 & book & 3629 & year \\ \hline
    20124 & year & 12770 & publisher & 3617 & publisher \\ \hline
    19786 & publisher & 12550 & url & 3444 & url \\ \hline
    19478 & url & 12220 & year & 3410 & book \\ \hline
    17360 & news & 11511 & news & 3273 & news \\ \hline
    16685 & accessdate & 11082 & united & 2968 & accessdate \\ \hline
    16401 & united & 10691 & accessdate & 2582 & world \\ \hline
  \end{tabularx}
  \caption{Example top~10 results and their frequencies from several different job
configurations. The top~10 do not change much one changes the parameters to make the job
less accurate.}
  \label{table:top10}
\end{table*}
\end{center}
